\speculake: Hiding Malware in Speculative Execution

### 2018-07-27 paper/architecture.tex
first cut at threat model

>>>>

\subsection{Threat Model}

We assume a scenario where an adversary wishes to hide or obscure the behavior
of a malicious program (malware) from an analyst attempting to reverse engineer
it. We note this is distinct from the goals of evading malware detection, where
maleware escapes classification by an anti-virus or other automated tool. While
we believe \speculake could also be used to make automated detection more
difficult, our main focus is on reverse engineer resistance, useful for evading
manual classification concerned with malware behavior. For anti-virus evasion,
we refer the reader to several existing techniques that are sufficient to defeat
existing anti-virus
systems~\cite{jana2012abusing,oberheide2009polypack,ropinjector,swinnen2014one}.

We assume the adversary is able to install binaries on the target machine
(e.g. via a trojan or remote exploitation), and the analyst is attempting to
determine what the malware will do using traditional debugging tools. We assume
the analyst may be aware that speculative execution is used to obfuscate
behavior, but does not have special-purpose hardware that allows introspection of
the CPU's speculative state. This assumption is realistic, as modern processors
do not allow developers or other users to directly interact with proprietary CPU
optimizations and features.

We further assume that the malware has a specific trigger that the analyst is
not privy to. In our examples, this trigger is often behavior exhibited by some
other (potentially benign) process running on the same system as the malware.
As the adversary is able to control when such a trigger is deployed (potentially
remotely), the analyst will not be able to observe or force this trigger to
happen at will. We emphasize that while this may also be true for existing
trigger-based malware, analysts can often reverse engineer the trigger out of
the malware, for example by observing control flow within the malware and using
adaptive fuzzers~\cite{afl} to generate inputs that explore other execution
paths of the binary. In contrast, \speculake malware's trigger depends on
behavior in seemingly unrelated binaries, influencing malware behavior through
channels invisible to the analyst. As with typical malware, the analyst may
attempt to reverse engineer the trigger to revela the malware's behavior, but we
will show (in Section~\ref{subsec:nested-spec}) how to defeat this type of
analysis.
